Component response rate variation drives stability in large complex systems

### 2018-06-03 notebook/ms.tex
Fix package name and edit links

@@

\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\linenumbers
\bibliographystyle{amnatnat}

@@

\includegraphics{ms_files/figure-latex/unnamed-chunk-4-1.pdf}

@@

\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
introducing variation in component response rate
(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
Eigenvalues plotted in the absence of \(Var(\gamma)\) where
\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
interaction strengths (\(\sigma^{2}\)) but clusters eigenvalues toward
the distribution's centre (-1, 0). Black and red elipses in both panels
show the circle centred on the distribution in panels a and b,
respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
of \(\Re(\lambda) < 0\) are 0.549 and 0.547 for a and b, respectively.

\includegraphics{ms_files/figure-latex/unnamed-chunk-7-1.pdf}

@@

Higher stability in systems with variation in \(\gamma\) can be observed
by sampling \(\gamma_{i}\) values from various distributions. I now
focus on a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\)
(see Supplementary Information for other distributions, which give
similar results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig.
1b), \(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
comparison of \(M\) before and after variation in component response
rate. Figure 2 shows a comparison of eigenvalue distributions given
\(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). As
expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}}, when
\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 12.649.
Uniform variation in \(\gamma\) leads to a non-uniform distribution of
eigenvalues, some of which are clustered tightly around the centre of
the distribution, but others of which are spread outside the former
radius of 12.649 (red circle Fig 2b). This larger radius occurs because
the addition of \(Var(\gamma)\) increases the realised \(\sigma\) of
\(M\). The clustering and spreading of eigenvalues introduced by
\(Var(\gamma)\) can destabilise previously stable systems or stabilise
systems that are otherwise unstable. But where systems are otherwise too
complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
can often lead to stability above
May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
threshold of \(\sigma\sqrt{SC} > 1\).

@@

\includegraphics{ms_files/figure-latex/unnamed-chunk-9-1.pdf}

@@

To investigate the effect of \(Var(\gamma)\) on system stability, I
simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
\(S\) ranging from 2-32. One million \(M\) were simulated for each
\(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
versus \(\gamma \sim \mathcal{U}(0, 2)\). I found that the number of
stable random systems was consistently higher given \(Var(\gamma)\) than
when \(\gamma = 1\) (Fig. 3), and that the difference between the
probabilities of observing a stable system increased with an increase in
\(S\); i.e., the potential for \(Var(\gamma)\) to drive stability
increased with system complexity. For the highest values of \(S\),
nearly all systems that were stable given \(Var(\gamma)\) would not have
been stable given \(\gamma = 1\) (see Supplementary Information for full
results). This suggests that the stability of large systems might be
dependent upon variation in the response rate of their individual
components, meaning that factors such as generation time (in ecological
networks), transaction speed (in economic networks), or communication
speed (in social networks) needs to be considered when investigating the
stability of complex systems.

It is important to point out that \(Var(\gamma)\) is not stabilising per
se; that is, adding variation in \(\gamma\) to a particular system \(M\)
does not necessarily increase the probability that the system will be
stable (see Supplementary Information). Rather, systems that are
observed to be stable are more likely to vary in \(\gamma\), and for
this \(Var(\gamma)\) to be critical to their stability. This is caused
by the shift in the distribution of eigenvalues that occurs by
introducing \(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in
all \(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\)
values. To further investigate the potential of \(Var(\gamma)\) to be
stabilising, I used a genetic algorithm (the space of possible
\(\gamma\) values was too large to search
exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
see Supplementary Information). For each of 10000 random \(M\), the
genetic algorithm initialised 1000 different sets of
\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis
was performed on \(M\) using each set of \(\gamma\) values, and the 20
sets with the lowest \(\max\left(\Re(\lambda)\right)\) each produced 50
offspring with subsequent mutation and crossover between the resulting
new population of 1000 \(\gamma\) sets. The genetic algorithm terminated
if a stable \(M\) was found, 20 generations occurred, or a convergence
criteria of minimum fitness increase between generations was satisfied.
Across \(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were
found that resulted in stable systems with probabilities that were
orders of magnitude higher than when \(\gamma = 1\) (see Supporting
Information), meaning that stability could often be achieved by
manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\)
elements. Hence, managing the response rates of system components in a
targetted way can potentially facilitate the stabilisation of complex
systems through a reduction in dimensionality.

I have focused broadly on random complex systems, but it is also
worthwhile to consider more restricted interactions such as those of
specific ecological
networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
include systems in which all interactions are negative (competitive
networks), positive (mutualist networks), or \(i\) and \(j\) pairs have
opposing signs (predator-prey networks). In general, competitive and
mutualist networks tend to be destabilising, and predator-prey network
tend to be
stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.
When \(Var(\gamma)\) is applied to each, the proportion of stable
competitive and predator-prey networks increases, but the proportion of
stable mutualist networks does not (see Supplementary Information).
Additionally, when each component of \(M\) is interpreted as a unique
species and given a random intrinsic growth
rate\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}},
feasibility is not increased by \(Var(\gamma)\), suggesting that
variation in species generation time might be unlikely to drive
stability in purely multi-species networks (see Supplementary
Information).

@@

@@

>>>>

\usepackage{lmodern}
\usepackage{setspace}
\setstretch{1}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={Component response rate variation drives stability in large complex systems},
            pdfauthor={A. Bradley Duthie},
            colorlinks=true,
            linkcolor=blue,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\bibliographystyle{amnatnat}

@@

\includegraphics{unnamed-chunk-4-1.pdf}

@@

\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
introducing variation in component response rate
(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
Eigenvalues plotted in the absence of \(Var(\gamma)\) where
\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
interaction strengths (\(\sigma^{2}\)) but clusters eigenvalues toward
the distribution's centre (-1, 0). Black and red elipses in both panels
show the circle centred on the distribution in panels a and b,
respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
of \(\Re(\lambda) < 0\) are 0.548 and 0.552 for a and b, respectively.

\includegraphics{unnamed-chunk-7-1.pdf}

@@

Higher stability in systems with variation in \(\gamma\) can be observed
by sampling \(\gamma_{i}\) values from various distributions. I now
focus on a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\)
(see \hyperlink{SIstart}{Supplementary Information} for other distributions, 
which give
similar results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig.
1b), \(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
comparison of \(M\) before and after variation in component response
rate. Figure 2 shows a comparison of eigenvalue distributions given
\(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). As
expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}}, when
\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 12.649.
Uniform variation in \(\gamma\) leads to a non-uniform distribution of
eigenvalues, some of which are clustered tightly around the centre of
the distribution, but others of which are spread outside the former
radius of 12.649 (red circle Fig 2b). This larger radius occurs because
the addition of \(Var(\gamma)\) increases the realised \(\sigma\) of
\(M\). The clustering and spreading of eigenvalues introduced by
\(Var(\gamma)\) can destabilise previously stable systems or stabilise
systems that are otherwise unstable. But where systems are otherwise too
complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
can often lead to stability above
May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
threshold of \(\sigma\sqrt{SC} > 1\).

@@

\includegraphics{unnamed-chunk-9-1.pdf}

@@

To investigate the effect of \(Var(\gamma)\) on system stability, I
simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
\(S\) ranging from 2-32. One million \(M\) were simulated 
for each \(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
versus \(\gamma \sim \mathcal{U}(0, 2)\). I found that the number of stable 
random systems was
consistently higher given \(Var(\gamma)\) than when \(\gamma = 1\) (Fig.
3), and that the difference between the probabilities of observing a
stable system increased with an increase in \(S\); i.e., the potential
for \(Var(\gamma)\) to drive stability increased with system complexity.
For the highest values of \(S\), nearly all systems that were stable
given \(Var(\gamma)\) would not have been stable given \(\gamma = 1\) (see
\hyperlink{SIstart}{Supplementary Information} for full results). This 
suggests that the
stability of large systems might be dependent upon variation in the
response rate of their individual components, meaning that factors such
as generation time (in ecological networks), transaction speed (in
economic networks), or communication speed (in social networks) needs to
be considered when investigating the stability of complex systems.

It is important to point out 
that \(Var(\gamma)\) is not stabilising per se; that is,
adding variation in \(\gamma\) to a particular system \(M\) does not
necessarily increase the probability that the system will be stable (see
\hyperlink{SIstart}{Supplementary Information}). Rather, systems that are 
observed to be stable are more likely to vary in \(\gamma\), and for this
\(Var(\gamma)\) to be critical to their stability. This is caused by the
shift in the distribution of eigenvalues that occurs by introducing
\(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in all
\(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\) values. To
further investigate the potential of \(Var(\gamma)\) to be stabilising,
I used a genetic algorithm (the space of possible \(\gamma\)
values was too large to search
exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
see \hyperlink{SIstart}{Supplementary Information}). For each of 10000 
random \(M\), the
genetic algorithm initialised 1000 different sets of
\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis was
performed on \(M\) using 
each set of \(\gamma\) values, and the 20 sets with the
lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 offspring with
subsequent mutation and crossover between the resulting new population
of 1000 \(\gamma\) sets. The genetic algorithm terminated if a stable
\(M\) was found, 20 generations occurred, or a convergence criteria of
minimum fitness increase between generations was satisfied. Across
\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
resulted in stable systems with probabilities that were orders of
magnitude higher than when \(\gamma = 1\) (see \hyperlink{SIstart}{Supplementary
Information}), meaning that stability could often be achieved by 
manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\) elements.
Hence, managing the response rates of system components in a targetted
way can potentially facilitate the stabilisation of complex systems 
through a reduction in dimensionality.

I have focused broadly on random complex systems, but it is also
worthwhile to consider more restricted interactions such as those of
specific ecological
networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or \(i\) and
\(j\) pairs have opposing signs (predator-prey networks). In
general, competitive and mutualist networks tend to be destabilising,
and predator-prey network tend to be
stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.
When \(Var(\gamma)\) is applied to each, the proportion of stable
competitive and predator-prey networks increases, but the proportion of
stable mutualist networks does not (see 
\hyperlink{SIstart}{Supplementary Information}).
Additionally, when each component of \(M\) is interpreted as a unique
species and given a random intrinsic growth
rate\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}},
feasibility is not increased by \(Var(\gamma)\), suggesting that
variation in species generation time might be unlikely to drive
stability in purely multi-species networks (see \hyperlink{SIstart}{Supplementary
Information}).

@@

\textbf{Acknowledgements}

I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
Conversations with L. Bussi\`ere and N. Bunnefeld improved the quality of
this work.

@@

\clearpage

\begin{center}
\hypertarget{SIstart}{\section{Supplementary Information}\label{SIstart}}
\end{center}

\textbf{This supplementary information supports the manuscript
``Component response rate variation drives stability in large complex
systems'' with all of the code required to recreate the analysis in the
main text, and with additional analyses to support its conclusions. All
text, code, and data underlying this manuscript are publicly available
on \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}
as part of the RandomMatrixStability package.}

The
\href{https://github.com/bradduthie/RandomMatrixStability}{RandomMatrixStability
package} includes all functions and tools for recreating the text, this
supplemental information, and running all code; additional documentation
is also provided for functions as part of the package. The
RandomMatrixStability package package is available on
\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}; to
download it, the
\href{https://cran.r-project.org/web/packages/devtools/index.html}{\texttt{devtools}
library} is needed.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{);}
\KeywordTok{library}\NormalTok{(devtools);}
\end{Highlighting}
\end{Shaded}

The code below installs the RandomMatrixStability package using
devtools.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install_github}\NormalTok{(}\StringTok{"bradduthie/RandomMatrixStability"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

While downloading this package is recommended, all relevant code is also
reproduced below with explanation, so it is possible to recreate all
analyses using only this Supplementary Information.

\section{Supplementary Information table of
contents}\label{supplemental-information-table-of-contents}

\begin{itemize}
\tightlist
\item
  \protect\hyperlink{Fig1}{Code and simulations underlying Fig. 1}
\item
  \protect\hyperlink{Fig2}{Code and simulations underlying Fig. 2}
\item
  \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
\item
  \protect\hyperlink{ecological}{Stability of ecological networks}

  \begin{itemize}
  \tightlist
  \item
    \protect\hyperlink{competition}{Competitor networks}
  \item
    \protect\hyperlink{mutualism}{Mutualist networks}
  \item
    \protect\hyperlink{pred-prey}{Predator-prey networks}
  \end{itemize}
\item
  \protect\hyperlink{connectance}{Different connectance (C) values}

  \begin{itemize}
  \tightlist
  \item
    \protect\hyperlink{connect3}{C = 0.3}
  \item
    \protect\hyperlink{connect5}{C = 0.5}
  \item
    \protect\hyperlink{connect7}{C = 0.7}
  \item
    \protect\hyperlink{connect9}{C = 0.9}
  \end{itemize}
\item
  \protect\hyperlink{gam_dist}{Different distributions of \(\gamma\)}
\item
  \protect\hyperlink{ga}{Genetic algorithm}
\item
  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
\end{itemize}

\hypertarget{Fig1}{\section{Code and simulations underlying Fig.
1}\label{Fig1}}

The sample \(M\) used for the eigenvalue distributions in Fig. 1 of the
text is available on
\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/sim_results/bi_gamma}{GitHub},
and was produced by running the following function.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{find_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) >=}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \KeywordTok{return}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1));}
            \NormalTok{break;}
        \NormalTok{\}}
        \KeywordTok{print}\NormalTok{(iters);}
        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
    \NormalTok{\}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The above \texttt{find\_bgamma} function terminates when a matrix \(M\)
is found that is not stable when all component response rates are set to
\(\gamma = 1\), but is stable when half of component response rates are
\(1.95\) and half are \(0.05\). The function is used to illustrate the
concept of how fast versus slow component responses can cause a system
to become stable. Simulations were run for \texttt{iter\ =\ 1000000},
but terminated once an acceptable \texttt{A0} and \texttt{A1} were
found. The code below plots the eigenvalue distributions of \texttt{A0}
and \texttt{A1} in panels \textbf{a} and \textbf{b}, respectively. The
plot itself can be recreated with the function and code below.

\includegraphics{unnamed-chunk-4-1.pdf}

To find out how frequently \(M\) was stable given that all
\(\gamma = 1\) versus \(\gamma = \{1.95, 0.05\}\), the function below
was created.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{stab_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
    \NormalTok{ress     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iters, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
    \NormalTok{A0_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
    \NormalTok{A1_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \NormalTok{ress[iters, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{A0_count       <-}\StringTok{ }\NormalTok{A0_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \NormalTok{ress[iters, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{A1_count       <-}\StringTok{ }\NormalTok{A1_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \NormalTok{\}}
        \KeywordTok{print}\NormalTok{(}\KeywordTok{c}\NormalTok{(iters, A0_count, A1_count));}
        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(ress);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The above functions produced the \texttt{bi\_pr\_st} data.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bi_pr_st <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"sim_results/bi_gamma/bi_pr_st.csv"}\NormalTok{);}
\NormalTok{pr_st    <-}\StringTok{ }\NormalTok{bi_pr_st[,-}\DecValTok{1}\NormalTok{];}
\end{Highlighting}
\end{Shaded}
